---
title: "Analyzing Building Permit Data"
output: 
  html_notebook: 
    code_folding: hide
    theme: simplex
---

# Preprocessing
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)

setwd("/Users/chaoguo/src/DurhamUpzoning/Data")
permit_data <- read.csv('All_Building_Permits.csv', header = TRUE, stringsAsFactors = FALSE)

# Some data wrangling
permit_data$ISSUE_DATE <- lubridate::ymd_hms(permit_data$ISSUE_DATE)
permit_data$Issue_year_month <- paste0(lubridate::year(permit_data$ISSUE_DATE), '-', lubridate::month(permit_data$ISSUE_DATE), '-', '01')
permit_data$Issue_year_month <- lubridate::ymd(permit_data$Issue_year_month)
```

# Percentage of permit application that were approved/issued
We want to see if there has been more permits that are related to building new residential homes after October 2019. The graph suggests an upward trend but we don't see a significant jump after October 2019.

We have also checked the percentage of these permits (divided by all permits), but its proportion is too high that most of the time, 100% permits are these permits.
```{r, fig.width = 10, fig.height = 8}
permit_status_by_time <- permit_data %>% group_by(PmtStatus, Issue_year_month) %>% summarise(cnt = n()) %>% ungroup() %>%
  tidyr::pivot_wider(id_cols = Issue_year_month, names_from = PmtStatus, values_from = cnt, names_prefix = 'PERMIT_TYPE_', values_fill = 0) %>%
  arrange(Issue_year_month)

permit_status_by_time$total_records <- rowSums(permit_status_by_time[, 2:10])

permit_status_by_time$Issued_OR_pending <- permit_status_by_time$`PERMIT_TYPE_CO Issued` +
  permit_status_by_time$`PERMIT_TYPE_CO Pending` +  # CO stands for "commercial/office"? Or "Conditional Overlays"?
  permit_status_by_time$PERMIT_TYPE_Complete +
  permit_status_by_time$`PERMIT_TYPE_Cond. CO Issued` +
  permit_status_by_time$PERMIT_TYPE_Issued

permit_status_by_time$Issued_OR_pending_perc <- permit_status_by_time$Issued_OR_pending / permit_status_by_time$total_records

p <- ggplot(permit_status_by_time[2:nrow(permit_status_by_time), ], aes(x=Issue_year_month, y=total_records)) +
  geom_line() + geom_point() +
  ggtitle('Number of Permits by Year/Month', subtitle = 'We are counting anything that is pending, issued, or complete') +
  scale_x_date(date_labels = "%b-%Y", date_breaks = '12 month') +
  xlab('Issued Year/Month') + ylab('Number of Records') +
  geom_vline(xintercept = as.numeric(as.Date("2019-10-01")), linetype=4) +  # zoning reforms implemented on October 1, 2019
  theme_linedraw()

print(p)
```

# Different types of activities by year
We make these plot by looking at different types of building types and building activities. The idea is that these two fields tell us the intention of the applicant.

## Adding/New a new modular unit
```{r, fig.width = 10, fig.height = 8}
n_by_year <- permit_data %>%
  filter(BLDB_ACTIVITY_1 %in% c('Addition','New'), BLD_Type %in% c('Modular Unit')) %>%
  group_by(Issue_year_month) %>% summarise(num_of_record = n())

ggplot(n_by_year[2:nrow(n_by_year), ], aes(x=Issue_year_month, y=num_of_record)) +
  geom_line() + geom_point() +
  ggtitle('Number of Permits by Year/Month', subtitle = 'Addition/New of Modular') +
  scale_x_date(date_labels = "%b-%Y", date_breaks = '12 month') +
  xlab('Issued Year/Month') + ylab('Number of Permits') +
  geom_vline(xintercept = as.numeric(as.Date("2019-10-01")), linetype=4) +  # zoning reforms implemented on October 1, 2019
  theme_linedraw()
```

## Adding/New a new SFH or a mobile home
```{r, fig.width = 10, fig.height = 8}
n_by_year <- permit_data %>% filter(BLDB_ACTIVITY_1 %in% c('Addition','New'),
                                    BLD_Type %in% c('Single Family House','Mobile Home') 
                                    )%>%
  group_by(Issue_year_month) %>% summarise(num_of_record = n())

ggplot(n_by_year[2:nrow(n_by_year), ], aes(x=Issue_year_month, y=num_of_record)) +
  geom_line() + geom_point() +
  ggtitle('Number of Permits by Year/Month', subtitle = 'Adding a Single Family House/Mobile Home') +
  scale_x_date(date_labels = "%b-%Y", date_breaks = '12 month') +
  xlab('Issued Year/Month') + ylab('Number of Permits') +
  geom_vline(xintercept = as.numeric(as.Date("2019-10-01")), linetype=4) +  # zoning reforms implemented on October 1, 2019
  theme_linedraw()
```

## Adding/New residential units that have more than 3 units
```{r, fig.width = 10, fig.height = 8}
n_by_year <- permit_data %>% filter(BLDB_ACTIVITY_1 %in% c('Addition','New'),
                                    BLD_Type %in% c('3-4 Residential Units','5+ Residential Units') 
                                    )%>%
  group_by(Issue_year_month) %>% summarise(num_of_record = n())

ggplot(n_by_year[2:nrow(n_by_year), ], aes(x=Issue_year_month, y=num_of_record)) +
  geom_line() + geom_point() +
  ggtitle('Number of Permits by Year/Month', subtitle = 'Addition/New of 3+ Residential Units') +
  scale_x_date(date_labels = "%b-%Y", date_breaks = '12 month') +
  xlab('Issued Year/Month') + ylab('Number of Permits') +
  geom_vline(xintercept = as.numeric(as.Date("2019-10-01")), linetype=4) +  # zoning reforms implemented on October 1, 2019
  theme_linedraw()
```